---
title: 'DATA 607: Assignment 2'
output:
  html_document:
    df_print: paged
---

For this assignment, I will be querying (contrived) data on recent popular movies. We will be using two data tables: `participants` and `ratings` to obtain relevant information. These tables were created via data definition language commands and are available in the .sql file in this repository. 

```{r}
library(tidyverse)
library(RSQLite)
library(RPostgres)
```


# Database Connection

```{r}

con = dbConnect(RPostgres::Postgres()
     , host='localhost'
     , port='5432'
     , dbname='deepikadilip'
     , user='deepikadilip')

tables <- dbListTables(con)

print(tables)

```

# Visualizating the Tables

## Participants Table
```{r}
res <- dbSendQuery(con, "SELECT * FROM participants")
tbl.participants = dbFetch(res)
dbClearResult(res)

head(tbl.participants)

```

## Ratings Table
```{r}
res <- dbSendQuery(con, "SELECT * FROM ratings")
tbl.ratings = dbFetch(res)
dbClearResult(res)

head(tbl.ratings)

```

# Exploratory Data Analysis
Now that we have the tables, we can do some exploratory data analysis.

Here is an overview of the number of ratings per movie:
```{r}
tbl.ratings %>% group_by(movie) %>% summarise(count = n()) %>% arrange( desc(count))

```

And how many ratings each participant provided:
```{r}
tbl.participants %>% left_join(tbl.ratings, by = "participant_id") %>% group_by(first_name, last_name) %>% summarise(count = n()) %>% arrange( desc(count))
```

As anticipated, some individuals didn't rate all the movies. Here, we'll produce a table of movies and the counts of the ratings totals. 

```{r}
tbl.merged = tbl.participants %>% left_join(tbl.ratings, by = "participant_id") 

tbl.ratings.N = tbl.ratings %>% group_by(movie) %>% summarise(n = n())

tbl.ratings.N = tbl.ratings.N %>% filter(n < 5)

tbl.ratings.N
```