---
title: 'DATA 607: Assignment 2'
output:
  html_document:
    df_print: paged
---

For this assignment, I will be querying (contrived) data on recent popular movies. We will be using two data tables: `participants` and `ratings` to obtain relevant information. These tables were created via data definition language commands and are available in the .sql file in this repository. 

```{r, echo=F, warning=F}
library(tidyverse)
library(RSQLite)
library(RPostgres)
```


# Database Connection

The first step is connecting to our local database. We have two tables available for analysis, `participants` and `ratings`:

```{r}

con = dbConnect(RPostgres::Postgres()
     , host='localhost'
     , port='5432'
     , dbname='deepikadilip'
     , user='deepikadilip')

tables <- dbListTables(con)

print(tables)

```

# Visualizating the Tables

## Participants Table

```{r}
res <- dbSendQuery(con, "SELECT * FROM participants")
tbl.participants = dbFetch(res)
dbClearResult(res)

head(tbl.participants)

```

## Ratings Table

```{r}
res <- dbSendQuery(con, "SELECT * FROM ratings")
tbl.ratings = dbFetch(res)
dbClearResult(res)

head(tbl.ratings)

```

# Exploratory Data Analysis
Now that we have the tables, we can do some exploratory data analysis.

* Here is an overview of the number of ratings per movie:

```{r}
tbl.ratings %>% group_by(movie) %>% summarise(count = n()) %>% arrange( desc(count)) 

```

* And how many ratings each participant provided:

```{r}
tbl.participants %>% left_join(tbl.ratings, by = "participant_id") %>% group_by(first_name, last_name) %>% summarise(count = n()) %>% arrange( desc(count))
```



# Missing Data

As anticipated, some individuals didn't rate all the movies. Here, we'll produce a table of movies and the counts of the ratings totals. This table will show all the movies that have fewer than 5 reviews (since there were five individuals). 

The first step is merging the participants and ratings tables using a left join:

```{r}
tbl.merged = tbl.participants %>% left_join(tbl.ratings, by = "participant_id") 

```

Afterwards, we'll tabulate the movies with less than 5 reviews:
```{r}

tbl.ratings.N = tbl.ratings %>% group_by(movie) %>% summarise(TOTAL = n())

tbl.ratings.N = tbl.ratings.N %>% filter(TOTAL < 5)

tbl.ratings.N
```


We'll also take a look at basic summary metrics for all the movies.

```{r}

tbl.ratings %>% group_by(movie) %>% summarise(MEAN = mean(rating), SD = sd(rating), TOTAL = n())

```

The last step is tabulate how many ratings each individual submitted, along with their average rating/standard deviation:

```{r}
tbl.merged %>% group_by(first_name, last_name) %>% summarise(n = n(), MEAN = mean(rating), SD = sd(rating))
```

# Takeaways

Based on this, we can conclude the following:

* Some will rate movies higher than others. Therefore, scaling the ratings per individual might be a better way to standardize the data.
* Not all movies have a sufficient number of ratings. For the purposes of determining the best movie, excluding those with only one rating (SPIDERMAN and LICORICE PIZZA) would be pratical.