---
title: 'Data 607: Assignment 3'
author: Deepika Dilip
output:
  html_document:
    df_print: paged
---

```{r, message=F, warning=F}
library(tidyverse)
library(ggplot2)
library(stringr)
```


# Problem 1

First we'll read in the dataset and then subset based on data/statistics strings

```{r}
df.college.majors = read.csv( url("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))

vec.majors = df.college.majors$Major[grep("DATA|STATISTICS", df.college.majors$Major)] 
print(vec.majors)
```


# Problem 2

To convert this to a properly formatted character string, we'll take the following steps:

1. Remove bracketed numbers from the string
2. Split the string based on quotation marks
3. Subset based on values that contain letters


```{r}
vec.text = c('[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"')

vec.text.char = gsub("(\\n\\[\\d+\\])|(^\\[\\d+\\])", "", vec.text)
vec.text.char = strsplit(vec.text.char, '\\"')
vec.text.char = unlist(vec.text.char)
vec.text.char = vec.text.char[grep("[a-z]", vec.text.char)]
print(vec.text.char)
```

# Problem 3


> 1.`(.)\1\1`

> 2. `(.)(.)\\2\\1`

> 3. `(..)\1`

> 4. `"(.).\\1.\\1"`

> 5. `"(.)(.)(.).*\\3\\2\\1"`

1. One character appears three times in a row
2. One character appears, followed by a second character twice, followed by the same character
3. Two characters (can be different) repeated consecutively (e.g. `dodo`)
4. An original character, followed by any character followed by the original character, followed by any character, and then the original character again.
5. Three original characters, followed by any number of characters, followed by the same characters in reversed order.


# Problem 4: 

**Construct regular expressions to match words that:** 

* Start and end with the same character.

```{r}
str_view("civic", "(.).*\\1$", match = T)
```

* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
str_view("church", "(..).*\\1", match = T)
```


* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
str_view("eleven", "(.).*\\1.*\\1", match = T)
```