---
title: "Assignment 10"
output:
  html_document:
    toc: true
    toc_float: yes
    theme: lumen
---
```{r warning = F, message=F}
library(topicmodels)
library(rvest)
library(tm)
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
library(ggplot2)
library(tidyverse)
library(janitor)
library(maps)
library(ggthemes)
library(tidytext)
library(textdata)
```

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Part I: Textbook Assignment

* Re-create and analyze primary code from the textbook.
* Provide citation to text book, using a standard citation syntax like APA or MLA.

For this section, I'll be using the `janeausten` package and replicating the example from the textbook:
```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

head(tidy_books, 5) %>% knitr::kable()
```

Essentially, the book is tokenized here, providing us with every single word of the book.  The next step will be joining our data with the `nrc` lexicon:

```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

Here we have a tabulation of terms (not verbatim). From here, we can tabulate the sentiments and visualize accordingly:
```{r}

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

The primary code also does a lexicon comparison:
```{r}
pride_prejudice <- tidy_books %>%
  filter(book == "Pride & Prejudice")

afinn <- pride_prejudice %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = linenumber %/% 80) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>%
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>%
    inner_join(get_sentiments("nrc") %>%
      filter(sentiment %in% c(
        "positive",
        "negative"
      ))) %>%
    mutate(method = "NRC")
) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

bind_rows(
  afinn,
  bing_and_nrc
) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y") +
  labs(title = "Sentiment Analysis of Pride and Prejudice")
```

Next, we'll visualize using a wordcloud:

```{r}

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 80))

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  reshape2::acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(
    colors = c("green", "purple"),
    max.words = 80
  )
```

The last step is aggregating sentences and attributing respective sentiments. We can see the ratio of negative to positive words per section:

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup() %>% 
  knitr::kable()
```

### References
Silge, Julia, and David Robinson, Text Mining with R: A Tidy Approach, O’Reilly Media, 2020. https://www.tidytextmining.com/


# Part II: Original Example
For this assignment, I'll be using tweets from the hashtag `#whyIdidntreport`. This hashtag trended around the confirmation of Justice Kavannaugh to the Supreme Court. I scraped the data and will be performing a sentiment analysis using the `nrc` lexicon. This will provide more details on the terms used. 

## Importing Tweets into a DataFrame
```{r message=F}
webpage <- read_html("tweet.html")
table <- html_table(webpage, fill = TRUE, header=TRUE)
# colnames(table) =  table[[1]]

data <-table[[1]]

colnames(data) = make_clean_names(colnames(data))
head(data, 2) %>% knitr::kable() %>% kableExtra::kable_styling()
```

## Subset to Original Tweets
There's quite a bit of information here, so we'll restrict this to the tweets
```{r message=F}
tweets <- subset(data, tweet_type == "Tweet")
```

## Text mining of tweets
We'll start by previewing some of the tweets
```{r message=F}
tweet_text <- Corpus(VectorSource(tweets$tweet))
(inspect(tweet_text[1:3]))
```

## Removing special characters
From here, we'll clean the text of spaces and unnecessary characters

```{r message=F, warning = F}
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
tweet_text2 <- tm_map(tweet_text, toSpace, "/")
tweet_text2 <- tm_map(tweet_text, toSpace, "@")
tweet_text2 <- tm_map(tweet_text, toSpace, "\\|")
```

## Text Cleaning
Here, we'll transform the tweets into corpus objects
```{r message=F, warning = F}
## Convert the text to lower case
docs <- tm_map(tweet_text2, content_transformer(tolower))
## Remove numbers
tweet_text2 <- tm_map(tweet_text2, removeNumbers)
## Remove english common stopwords
tweet_text2 <- tm_map(tweet_text2, removeWords, stopwords("english"))
## specify your stopwords as a character vector
tweet_text2 <- tm_map(tweet_text2, removeWords, c("I", "my", "because")) 
## Remove punctuations
tweet_text2 <- tm_map(tweet_text2, removePunctuation)
## Eliminate extra white spaces
tweet_text2 <- tm_map(tweet_text2, stripWhitespace)
```

## Topic Analysis
We can also separate the tweets by topic:
```{r message=F}
dtm <- DocumentTermMatrix(tweet_text2)
dtm <- removeSparseTerms(dtm, 0.99)
ldaOut <-LDA(dtm, k = 4)
topics <-terms(ldaOut,6)
head(topics, 2) %>% knitr::kable()
```

## Document Matrix
The next step is a document matrix which sorts words by frequency
```{r message=F}
tweets_matrix <- TermDocumentMatrix(tweet_text2)
m <- as.matrix(tweets_matrix)
v <- sort(rowSums(m),decreasing=TRUE)
doc_matrix <- data.frame(word = names(v),freq=v)
```

## Building the Word Chart
Here we'll plot the word frequencies:

```{r}
graph = ggplot(data = doc_matrix[8:30,], aes(reorder(word, -freq), freq)) + geom_bar(stat = "identity", fill = "lightblue", color = "black") + labs( title ="Most frequent words", x = "Word", y = "Word frequencies") + theme_minimal() + theme(axis.text.x=element_text(angle=90))
plot(graph)
```

## Building the Word Cloud

We can preview term frequency using a wordcloud:

```{r}
wordcloud <- wordcloud(words = setdiff(doc_matrix$word, c("whyididntreport", "because")), freq = doc_matrix$freq, min.freq = 1,
          max.words=50, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

## Sentiment Analysis
Here we'll use the `nrc` lexicon and visualize the top sentiments:

```{r, message = FALSE}

nrc_sent <- get_sentiments("nrc")
# bing_sent <-get_sentiments("bing")
#affin_sent <-get_sentiments("afinn")

tweet_nrc_sent <- doc_matrix %>%
  inner_join(nrc_sent, by=c("word")) 

#tweet_bing_sent <- doc_matrix %>% inner_join(bing_sent, by=c("word"))

# tweet_affin_sent <- doc_matrix %>% inner_join(affin_sent, by=c("word"))

tweet_nrc_sum <- tweet_nrc_sent %>%
  select(sentiment, freq) %>% 
  group_by(sentiment) %>%
  rename (count = freq) %>%
  summarise(n = sum(count)) %>% arrange(n) %>% filter(sentiment != "negative" & sentiment != "positive")

 tweet_nrc_sum %>% ggplot(aes(reorder(sentiment, -n), n, fill = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() + theme_clean()
```
