---
title: "Assignment 5"
author: "Deepika Dilip"
date: "2/26/2022"
output: 
  html_document:
    theme: cosmo
    toc: true
---


# Import Libraries

```{r setup, warning=F, message = F}
library(data.table)
library(ggplot2)
library(janitor)
library(tidyverse)
library(knitr)
library(XML)
library(kableExtra)
library(htmltools)
library(rjson)

dir.home = '~/git/CUNY.MDS/DATA607/'
setwd(dir.home)
```

# Data Preview

For reference, here's the book data, we'll be working with

```{r}
fread('books.csv') %>% kable() %>% kable_styling(bootstrap_options = "basic")
```

# XML File

I started by manually creating an XML file:
```{r, echo = F}
vec.xml = c('<?xml version="1.0" encoding="UTF-8"?>
<bookstore>
  <book>
    <title>And Then There Were None</title>
    <Author>Agatha Christie</Author>
    <Attribute.1>Thrilling</Attribute.1>
    <Attribute.2>Suspenseful</Attribute.2>
  </book>
  <book>
    <title>Dreamland</title>
    <Author>Sam Quinones</Author>
    <Attribute.1>Informative</Attribute.1>
    <Attribute.2>Saddening</Attribute.2>
    <Attribute.3>Detailed</Attribute.3>
  </book>
  <book>
    <title>The Great Gatsby</title>
    <Author>Scott Fitzgerald</Author>
    <Attribute.1>Thoughtful</Attribute.1>
    <Attribute.2>Tragic</Attribute.2>
  </book>
</bookstore>')
writeLines(vec.xml)
```

From here, we'll use the built-in R command to convert the data

```{r}
xmlToDataFrame(vec.xml) %>% kable() %>% kable_styling(bootstrap_options = "basic")
```

# HTML File

Similar to the XML file, I manually created an HTML file:
```{r, echo = F}
vec.html = c('<html>
<table>
 <tr>
  <td>Title</td>
  <td>Author</td>
  <td>Attribute.1</td>
  <td>Attribute.2</td>
  <td>Attribute.3</td>
 </tr>
 <tr>
  <td>And Then There Were None</td>
  <td>Agatha Christie</td>
  <td>Thrilling</td>
  <td>Suspenseful</td>
 </tr>
 <tr>
  <td>Dreamland</td>
  <td>Sam Quinones</span></td>
  <td>Informative</td>
  <td>Saddening</td>
  <td>Detailed</td>
 </tr>
 <tr>
  <td>The Great Gatsby</td>
  <td>Scott Fitzgerald</td>
  <td>Thoughtful</td>
 </tr>
</table>
</html>')
writeLines(vec.html)
```

From here, we'll use the `xml` package to parse the file:

```{r,  results='asis'}
html.tbl = htmlParse(vec.html) %>% readHTMLTable()
html.tbl$`NULL` %>% kable() %>%  kable_styling(bootstrap_options = "basic")
```


# JSON File
I manually created a JSON file:
```{r, echo = F}
vec.json = c('[
 {
   "Title": "And Then There Were None",
   "Author": "Agatha Christie",
   "Attribute.1": "Thrilling",
   "Attribute.2": "Suspenseful",
   "Attribute.3": ""
 },
 {
   "Title": "Dreamland",
   "Author": "Sam Quinones",
   "Attribute.1": "Informative",
   "Attribute.2": "Saddening",
   "Attribute.3": "Detailed"
 },
 {
   "Title": "The Great Gatsby",
   "Author": "Scott Fitzgerald",
   "Attribute.1": "Thoughtful",
   "Attribute.2": "Tragic",
   "Attribute.3": ""
 }
]')
writeLines(vec.json)
```


```{r}
lst.json = fromJSON(vec.json)
do.call(rbind, lst.json) %>% kable() %>%  kable_styling(bootstrap_options = "basic")
```


# Takeaways

* The HTML and XML formats populated the table with `NA` values
* The JSON formats preserved the blank cells

