---
title: "Project 2, Dataset 1"
author: "Deepika Dilip"
date: "3/12/2022"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(ggthemes)
library(plotly)
library(lubridate)
library(kableExtra)
library(tidyverse)
```


# Overview

For this analysis, we'll be looking at COVID-19 cases and deaths by state over time. We'll be using the CDC dataset. For our analysis, we will be doing the following:

1. Compare and comfirm/probable cases by season and state
2. Compare and comfirm/probable deaths by season and state
3. Compare the time gap between data creation and submission by state

## Step 1: CSV
First, we'll read the dataset in
```{r cars}
setwd('~/git/CUNY.MDS/DATA607/')
dataset.1 = as.data.table(read.csv(url('https://data.cdc.gov/resource/9mfq-cb36.csv')))
head(dataset.1) %>% kable() %>% kable_styling( full_width = F)
```

Right off the bat, we notice this dataset has a wide structure--which is great for the purposes of exporting the CSV file! We do need to manipulate the date stamps though and change a couple of these variables to booleans:

```{r}
dataset.1[ , `:=` (submission_date = as_date(submission_date), created_at = as_date(created_at))]
dataset.1[consent_cases == 'Agree', consent_cases.boolean := T]
dataset.1[consent_cases == 'Not agree', consent_cases.boolean := F]
dataset.1[!(consent_cases %in%  c('Agree', 'Not agree')), consent_cases.boolean := NA]

dataset.1[consent_deaths == 'Agree', consent_deaths.boolean := T]
dataset.1[consent_deaths == 'Not agree', consent_deaths.boolean := F]
dataset.1[!(consent_deaths %in%  c('Agree', 'Not agree')), consent_deaths.boolean := NA]

head(dataset.1) %>% kable() %>% kable_styling( full_width = F)

```


We also need to create the `season` variable. To do this, we'll create a function to do just that:

```{r}

getSeason = function(x) {
 winter = list(lubridate::interval(ymd("19721221"),ymd("19721231")), lubridate::interval(ymd("19720101"),ymd("19720319")))
  #winter <- lubridate::interval(ymd("19720101"),ymd("19720319"))
  spring <- lubridate::interval(ymd("19720320"),ymd("19720619"))
  summer <- lubridate::interval(ymd("19720620"),ymd("19720921"))
  fall <- lubridate::interval(ymd("19720922"),ymd("19721220"))
  x = as_date(x)
  year(x) = 1972
  if (x %within% winter) {
    return("winter")
  } else if (x %within% spring) {
    return("spring")
  } else if (x %within% summer) {
    return("summer")
  } else if (x %within% fall) {
    return("fall")
  } else {
    return(NA)
  }
}

dataset.1[, season := sapply(submission_date, getSeason)]
dataset.1[, table(season)] %>% kable() %>% kable_styling( full_width = F)

```


From here we can export
```{r}
write.csv(dataset.1, 'dataset1.csv', row.names = F)
```

# Step 2: Analysis

*Season and State Correlations*

From our manufactured variables, we can produce the tables of interest:
```{r}
dataset.1[order(state), list(prob_cases_total = sum(prob_cases, na.rm = T), prob_deaths_total = sum(prob_death, na.rm = T)), by = "state"]%>% kable() %>% kable_styling( full_width = F)
dataset.1[order(season), list(prob_cases_total = sum(prob_cases, na.rm = T), prob_deaths_total = sum(prob_death, na.rm = T)), by = "season"] %>% kable() %>% kable_styling( full_width = F)
```

### Time Gap

The time gap would involve creating a new variable. We can aggregate it by mean time gap per state:

```{r}
dataset.1[, time.gap := difftime(created_at, submission_date, units = "days")]
time.lag = dataset.1[order(state), list(mean_time_gap = mean(time.gap, na.rm = T)), by = "state"]  

time.lag %>% kable() %>% kable_styling( full_width = F)

plt.time.lag = ggplot(time.lag, aes(x = state, y = mean_time_gap)) + geom_bar(fill = 'lightpink', stat = "identity") + theme_minimal()  + labs(x = "State", y = "Mean Time Gap", title = "Time Gap per State") + theme(axis.text.x =  element_text(angle = 90, vjust = 0.5, hjust=1))

ggplotly(plt.time.lag)
```

The states with the highest time lag values:
```{r}
time.lag[order(mean_time_gap, decreasing = T),][1:10] %>% kable() %>% kable_styling( full_width = F)
```

```{r}
dataset.1[, year := year(submission_date)]
dataset.1[, season := factor(season, c("winter", "spring", "summer", "fall"))]

season.cases = dataset.1[,list(prob_cases_total = sum(prob_cases, na.rm = T), prob_deaths_total = sum(prob_death, na.rm = T)), by = c("year", "season")]
season.cases = season.cases[order(year, season)]
```

### Season Plots

```{r}
season.cases[, time.period := paste0(year,"_", season)]
plt.seasons.cases = ggplot(season.cases, aes(x = season, y = prob_cases_total)) + geom_bar(fill = 'darkblue', stat = "identity") + theme_minimal()  + labs(x = "Season/Year", y = "Probable Cases (Total)", title = "Probable Cases per Season/Year") + theme(axis.text.x =  element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(~year)
plt.seasons.deaths = ggplot(season.cases, aes(x = season, y = prob_deaths_total)) + geom_bar(fill = 'seagreen', stat = "identity") + theme_minimal()  + labs(x = "Season/Year", y = "Probable Deaths (Total)", title = "Probable Deaths per Season/Year") + theme(axis.text.x =  element_text(angle = 90, vjust = 0.5, hjust=1)) + facet_wrap(~year)

ggplotly(plt.seasons.cases)
ggplotly(plt.seasons.deaths)

```


# Key Takeaways

Based on our analysis, we can conclude the following:

* Fall to winter indicated the higher rate of increase for probable cases
* The trend was also observed in probable deaths BUT the trend oscillated a bit. This is most likely due to improved surveillance; as a result the death count is less likely to be probable and more likely to be confirmed.
* ME, IN, NE, CA, MS, and MD had some of the states with the highest time gaps between probable cases and deaths.
