---
title: "Project 3"
output:
  html_document:
    toc: true
---

# Acute Myeloid Leukemia

In this dataset, I'll be using data from `cBioPortal` to analyze cases of acute myeloid leukemia.


```{r warning = F, message = F}
setwd("~/git/CUNY.MDS/DATA607/")
library(data.table)
library(snakecase)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggpubr)
library(broom)
```

## Feature Selection
For our purposes, we'll be using survival data, alongside mutation data. We can subset the dataset for variables of interest:

```{r warning = F, message = F}
dataset3 = fread("aml_ohsu_2018_clinical_data.tsv")
names(dataset3) = snakecase::to_mixed_case(names(dataset3))

dataset3 = dataset3[, .(Patient_Id, Sample_Id, Overall_Survival_Months, Overall_Survival_Status, Tp_53_Mutation, Dnmt_3_A_Mutation, Eln_2017_Risk_Classification, Npm_1_Consensus_Call, Cumulative_Treatment_Regimens)]


```


# CSV Conversion:
For the purposes of this assignment, we'll be using risk stratification for converting the dataset to wide form and populating the values with the respective mutations:

Thei first step is consolidating risk categories:

```{r warning = F, message = F}
dataset3[, .N, by = "Eln_2017_Risk_Classification"] %>% kable %>% kable_styling( full_width = F) 
```

This is a bit messy, so we'l consolidate: 

```{r}
dataset3[grepl("favorable", Eln_2017_Risk_Classification, ignore.case = T), Risk.Classification := "Good"]
dataset3[grepl("Adverse", Eln_2017_Risk_Classification, ignore.case = T), Risk.Classification := "Poor"]
dataset3[grepl("Intermediate", Eln_2017_Risk_Classification, ignore.case = T), Risk.Classification := "Intermediate"]
dataset3[is.na(Risk.Classification), Risk.Classification := "Unknown"]
dataset3[, Risk.Classification := factor(Risk.Classification, levels = c("Poor", "Intermediate", "Good"))]
dataset3[, .N, by = "Risk.Classification"] %>% kable %>% kable_styling( full_width = F, position = "left") 
```

Most patients here have one sample but some have two. For the sake of the export, we'll parse the treatment regimens into wide form.

```{r}
treatment = dataset3[,.(Patient_Id, Cumulative_Treatment_Regimens)]
lst.split = strsplit(treatment$Cumulative_Treatment_Regimens, split = "\\|")

lst.split.new = list()
for (i in 1:length(lst.split)) {
  vec = lst.split[[i]]
  vec = unlist(vec)
  if (length(vec) < 13) {
    Y = 13 - length(vec)
    na.vec = rep(NA, Y)
    vec = c(vec, na.vec)
  }
  lst.split.new[[i]] = vec
}


mat.treatments = do.call(rbind, lst.split.new)
colnames(mat.treatments) = paste0("treatment_", 1:ncol(mat.treatments))
rownames(mat.treatments) = treatment$Patient_Id

dt.treatment = as.data.table(mat.treatments, keep.rownames = "Patient_Id")
```

We can merge this data.table back and export:
```{r}
dt.wide = merge(dataset3[, .(Patient_Id, Sample_Id, Overall_Survival_Months, Overall_Survival_Status, Tp_53_Mutation, Dnmt_3_A_Mutation, Risk.Classification, Npm_1_Consensus_Call)], dt.treatment, by = "Patient_Id")

head(dt.wide) %>% kable %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "left") 
write.csv(dt.wide, file = "dataset3.csv")
```

# Data Analysis:

The first step is making the mutation data boolean expressions. We'll also do the same with survival:


```{r}
dataset3[is.na(Tp_53_Mutation), TP53 := FALSE]
dataset3[!is.na(Tp_53_Mutation), TP53 := TRUE]

dataset3[is.na(Npm_1_Consensus_Call), NPM1 := FALSE]
dataset3[!is.na(Npm_1_Consensus_Call), NPM1 := TRUE]

dataset3[is.na(Dnmt_3_A_Mutation), DNMT3A := FALSE]
dataset3[!is.na(Dnmt_3_A_Mutation), DNMT3A := TRUE]

dataset3[grepl("deceased", Overall_Survival_Status, ignore.case = T), SURVIVAL := 1]
dataset3[grepl("alive", Overall_Survival_Status, ignore.case = T), SURVIVAL := 0]

```

From here, we'll fit Cox-Proportional Hazards models for : 1) TP53 mutations 2) DNMT3A mutations 3) Interactions between TP53/DNMT3A 4) NPM1 mutations

```{r}

tbl = rbind(tidy(coxph(Surv(Overall_Survival_Months, SURVIVAL) ~ TP53, data=dataset3)), 
      tidy(coxph(Surv(Overall_Survival_Months, SURVIVAL) ~ DNMT3A, data=dataset3)),
      tidy(coxph(Surv(Overall_Survival_Months, SURVIVAL) ~ NPM1, data=dataset3)))

tbl %>% kable() %>%  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position= "left") 

tidy(coxph(Surv(Overall_Survival_Months, SURVIVAL) ~ TP53*DNMT3A, data=dataset3)) %>% kable() %>%  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F, position = "left") 

```


We'll visualize these results as well:

```{r}

plot0 = ggsurvplot(
    fit = survfit(Surv(Overall_Survival_Months, SURVIVAL) ~ factor(Risk.Classification), data = dataset3), 
    xlab = "Days", 
    ylab = "Overall survival probability (TP53)", ggtheme = theme_minimal()) 

plot1 = ggsurvplot(
    fit = survfit(Surv(Overall_Survival_Months, SURVIVAL) ~ TP53, data = dataset3), 
    xlab = "Days", 
    ylab = "Overall survival probability (TP53)", ggtheme = theme_minimal()) 

plot2 = ggsurvplot(
    fit = survfit(Surv(Overall_Survival_Months, SURVIVAL) ~ DNMT3A, data = dataset3), 
    xlab = "Days", 
    ylab = "Overall survival probability (DNMT3A)", ggtheme = theme_minimal()) 


plot3 = ggsurvplot(
    fit = survfit(Surv(Overall_Survival_Months, SURVIVAL) ~ NPM1, data = dataset3), 
    xlab = "Days", 
    ylab = "Overall survival probability (NPM1)", ggtheme = theme_minimal()) 


plot0
plot1
plot2
plot3

```

# Conclusions

Based on our findings, we can conclude the following:

* TP53 mutations are statistically significantly associated with worse survival. This makes sense since TP53 is an oncogene. 
* DNMT3A is also associated with worse survival. This is most likely correlated with epigenetic markers that influence gene expression.
* Interaction with both mutations has an increased negative effect on prognosis.
* We most likely need to parse the risk classification in more detail because the model wasn't indicative.

